
---
title: "Data Visualization (incarceration)"

date: "2/24/2022"
author: "Sean"
output: html_document
---
# Data Visualization (incarceration)

### Introduction 
To know the population of each races in jail at Autauga County since 1970 to 2018, and 
the proportion of different races between 1990 to 2018. I created variables to compute
proportion of different races. I uses the sum of totoal people from 15 to 64 years old
in jail every years, and sum of total population from 15 to 64 years old of different 
races. Than I divided those two data, I got the proportion of differentraces in jail.

|Racial population | Proportion|
|----|----|
|Asian American / Pacific Islander Population Count:| 0.93%|
|Black Population Count:| 18.15%|
|Latinx Population Count:| 1.82%|
|Native American Population Count:| 0.42%|
|White Population Count:| 78.68%|

```{r echo=FALSE, warning=FALSE}
incarceration <- read.csv(file = 'https://raw.githubusercontent.com/vera-institute/incarceration-trends/master/incarceration_trends.csv')
head(incarceration)
View(incarceration)


library(dplyr)
sum_1970_2018 <- with(incarceration, sum(total_pop[fips == '1001']))
AL <- filter(incarceration, fips == '1001')
annual_pop_AL <- AL[ , c(6)] 
sum_1990_2018 <- with(incarceration, sum(total_pop[fips == '1001' & year > '1989']))
sum_1990_2018_15to64 <- with(incarceration, sum(total_pop_15to64[fips == '1001' & year > '1989']))
sum_aapi_1990_2018_15to64 <- with(incarceration, sum(aapi_pop_15to64[fips == '1001' & year > '1989']))
sum_black_1990_2018_15to64 <- with(incarceration, sum(black_pop_15to64[fips == '1001' & year > '1989']))
sum_latinx_1990_2018_15to64 <- with(incarceration, sum(latinx_pop_15to64[fips == '1001' & year > '1989']))
sum_native_1990_2018_15to64 <- with(incarceration, sum(native_pop_15to64[fips == '1001' & year > '1989']))
sum_white_1990_2018_15to64 <- with(incarceration, sum(white_pop_15to64[fips == '1001' & year > '1989']))

prop_aapi_1990_2018 <- (sum_aapi_1990_2018_15to64 / sum_1990_2018_15to64)
prop_black_1990_2018 <- (sum_black_1990_2018_15to64 / sum_1990_2018_15to64)
prop_latinx_1990_2018 <- (sum_latinx_1990_2018_15to64 / sum_1990_2018_15to64)
prop_native_1990_2018 <- (sum_native_1990_2018_15to64 / sum_1990_2018_15to64)
prop_white_1990_2018 <- (sum_white_1990_2018_15to64 / sum_1990_2018_15to64)
```

### Trends over time charts
So in chart one, I use a line plot to show how does population changes annually
between different races, and I use different colors to represent different races.
In the chart, the blue line shows the population from 15 to 64 in jail, white line
represent white people, black line represents black people, yellow line represents
Asian and pacific islander, brown represent latinx, gray represent native Americans.
Why I chose line plot to show the data is because, through the chart we can see 
clrealy, since 1990, the population of every race are increases, Whites are the 
largest racial group in prison.

```{r echo=FALSE, warning=FALSE}
library(ggplot2)
incarceration <- read.csv(file = 'https://raw.githubusercontent.com/vera-institute/incarceration-trends/master/incarceration_trends.csv')
head(incarceration)
View(incarceration)
sum_1970_2018 <- with(incarceration, sum(total_pop[fips == '1001']))
AL <- filter(incarceration, fips == '1001')
annual_pop_AL <- AL[ , c(6)] 
  p_line=ggplot(AL)+
    geom_line(aes(x=AL[,2],y=AL[,6]),color="red")+
    geom_line(aes(x=AL[,2],y=AL[,7]),color="blue")+
    geom_line(aes(x=AL[,2],y=AL[,10]),color="yellow")+
    geom_line(aes(x=AL[,2],y=AL[,11]),color="black")+
    geom_line(aes(x=AL[,2],y=AL[,12]),color="brown")+
    geom_line(aes(x=AL[,2],y=AL[,13]),color="gray")+
    geom_line(aes(x=AL[,2],y=AL[,14]),color="white")
  p_line
  p_line +xlab("Years")+
    ylab("Population in jail")+
    ggtitle('Population changes in jail') 
```

### Comparison
In chart two, I also use a line plot to show the comparison of populasion in 
jail between Asian& Pacific Islander American and native Americans. Because
I found that in chart one, the population between these two races are close.
So I do the chart two to compare this. Yellow represent Asian and Islander 
Americans, gray represent Native Americans. Through the graph we know that
the Asian and Islander Americans keep increase, and the Native Americans increase
in 1990 to 2000 and after 2000 it is relatively stable. The data is change around
150 every year.
```{r echo=FALSE, warning=FALSE}
  p_line_2=ggplot(AL)+
    geom_line(aes(x=AL[,2],y=AL[,10]),color="yellow")+
    geom_line(aes(x=AL[,2],y=AL[,13]),color="gray")
  p_line_2
  p_line_2 +xlab("Years")+
    ylab("Population in jail")+
    ggtitle('Comparison of prison population of Asian&Islander and Native') 
```






### Mapping

```{r echo=FALSE, warning=FALSE}
  library(maps)
  library(tmap) 
  library(leaflet) 
  library(ggplot2) 
  library(dplyr)
  library(tidyverse)
  library(ggmap)
  
map <- get_stamenmap(
    bbox = c(left = -87.4141, bottom = 32.0640, right = -85.9172, top = 32.9499),
    maptype = "toner",
    zoom = 10
  )
  
ggmap(map)
```


